---
title: "Clusters: Mouse snSeq"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(Seurat)
library(dittoSeq)
library(scCustomize)
library(DT)
```

###
Here you'll find:

* SCTransform normalization is performed, and marker genes are identified using the PrepSCTFindMarkers function.

* Top marker genes for each cluster are visualized and saved. 

* Literature-based marker genes are checked

###


# Read data

Data formated in a seurat object by `2_SeuratPipeline.qmd` file

```{r}
seurat <- readRDS( here::here("output/rds/3_seurat_annotation.RDS"))
Idents(seurat) <- "Level_2"
```


# Cluster markers

```{r}
options(future.globals.maxSize = 3e+09)
seurat[["RNA"]] <- split(seurat[["RNA"]], f = seurat$id)
seurat <- SCTransform(seurat, conserve.memory = T, ncells = 1000)
seurat <- PrepSCTFindMarkers(seurat)
```



```{r eval=T}

# find markers for every cluster compared to all remaining cells
# this step takes long time
Idents(seurat) <- "Level_4"
seurat.markers <- FindAllMarkers(seurat, 
                                 only.pos = T, 
                                 min.pct = 0.25, 
                                 logfc.threshold = 0.25)


top2 <- seurat.markers %>%
    group_by(cluster) %>%
    slice_head(n = 4)

DoHeatmap(seurat, features = top2$gene, group.by = "Level_2") + NoLegend()

DotPlot(seurat, features = unique(top2$gene), cluster.idents = T, group.by = "Level_4", ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

scCustomize::Clustered_DotPlot(seurat_object = seurat, 
                               features = unique(top2$gene),  x_lab_rotate = 90,
                               group.by = "Level_4", flip = F
                               ) 

openxlsx::write.xlsx(seurat.markers, here::here("output/tables/4_markers_Level_4.xlsx"))
saveRDS(seurat, here::here("output/rds/4_seurat_SCTnorm.RDS"))

```



```{r}
htmltools::tagList(
lapply(unique(seurat.markers$cluster), function(i){
  datatable(seurat.markers[which(seurat.markers$cluster == i),])
  })
)
```

Using FindConservedMarkers 

```{r}
# Idents(seurat) <- "Level_4"
# # Create function to get conserved markers for any given cluster
# get_conserved <- function(cluster){
#         FindConservedMarkers(seurat,
#                              ident.1 = cluster,
#                              grouping.var = "id",
#                              only.pos = TRUE) %>%
#                 rownames_to_column(var = "gene") %>%
#                 cbind(cluster_id = cluster, .)
# }
# 
# # Iterate function across desired clusters
# conserved_markers <- map_dfr(unique(seurat$Level_4), get_conserved)
# 
# top2.conserved <- conserved_markers %>%
#     group_by(cluster_id) %>% 
#   slice_head(n = 2)
# 
# DoHeatmap(seurat, features = top2.conserved$gene) + NoLegend()
# 
# openxlsx::write.xlsx(conserved_markers, here::here("output/tables/3_conserved_markers_Level_4_2024-04-23.RDS"))
```


# TSNE plot

```{r}
DimPlot(seurat, 
        label = F, reduction = "tsne", 
        cols = scCustomize::DiscretePalette_scCustomize(num_colors = 40, palette = "ditto_seq"),
        group.by  = "Level_2") & NoAxes()  +
  theme(legend.position = "top")


DimPlot(seurat, 
        label = T, reduction = "tsne", 
        cols = scCustomize::DiscretePalette_scCustomize(num_colors = 40, palette = "ditto_seq"),
        group.by  = "Level_4") & NoAxes()  +
  theme(legend.position = "top")
```


# Literature based markers

Ependymal cells reference: https://www.cell.com/cell/pdf/S0092-8674(18)30395-7.pdf 
Peri https://www.nature.com/articles/s41586-021-04369-3#MOESM2

```{r}
markers <- c(
  "P2ry12", "Sall1", "Ctss", "C1qb", "Cx3cr1", # Microglia 
  "Folr1", "Prlr", "Ttr",  # Choroid plexus
  "Ccr2", 
  "Foxj1", "Acta2", "Tuba1a", "Ccdc153", # Ependymal cells
  "Cntn1", "Tnr", "Vcan", "Pdgfra", # OPC
  "Plp1", "Mog", "Mbp", "Mal", # Oligodendrocyte
  "Gfap", "Aqp4", "Gja1", "Slc1a3", "Clu", "Slc1a2", # Astrocytes
  "Bsg", "Flt1", "Itm2a", "Ifitm1", "Apold1",  "Pecam1", # Endothelial
  "Snap25",  "Rbfox1", "Grin1", # Neuron 
  "Dcx", "Hs3st2", #  Excitatory Neuron 
  "Gad1", "Gad2", # Interneurons 
  "Grin2b",
  "Ptprc", "Runx1", # Immune
  "Cemip","Bicc1","Flrt2", "Lama2" # Perivascular_Fibroblast
) %>% 
  unique()

DotPlot(seurat, features = markers, cluster.idents = T, group.by = "Level_2") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

scCustomize::Clustered_DotPlot(seurat_object = seurat, 
                               features = markers,  x_lab_rotate = 90,
                               group.by = "Level_4", k = 10, flip = F
                               ) 
```


Dawn's markers from Fei's lab

```{r}
markers <- c(
 
#Excitatory neurons, amygdala
"Nrgn", "C1ql3", "Olfm1", "Rtn1", "Cck", "Mef2c", "Calm2", "Chgb", "Ywhah", "Ctxn1",
#Excitatory neurons, cerebral cortex
"Meg3", "Snhg11", "Rian", "R3hdm1", "Kcnq1ot1", "Grin2a", "Snhg14", "Grin2b", "Syt7", "Ahi1",
"Dcx", "Hs3st2", #  Excitatory Neuron 
"Gad1", "Gad2", # Interneurons
#Astrocytes
"Slc1a3", "Apoe", "Aldoc", "Plpp3", "Atp1a2", "Slc1a2", "Ntsr2", "Gja1", "Mt1", "Bcan",
#R-LM border Cck interneurons
"Tnr", "Mbp",
#Mature oligodendrocytes
"Plp1", "Mal", "Cnp", "Mag", "Trf", "Mobp", "Cldn11", "Mbp", "Cryab", "Ermn",
"Cntn1", "Tnr", "Vcan", "Pdgfra", # OPC
#Dz medium spiny neurons", "striatum
"Penk", "Pde10a", "Gpr88", "Tac1", "Rgs9", "Adcy5", "Ppp1r1b", "Pde1b", "Rasd2", "Scn4b",
#Microglia", "activated
"Hexb", "C1qa", "C1qb", "Ctss", "C1qc", "Cx3cr1", "Csf1r", "P2ry12", "Ctsd", "Tyrobp",
#Choroid plexus epithelial cells
"Ttr", "Enpp2", "Ecrg4", "Folr1", "2900040C04Rik", "Prlr", "Igfbp2", "Sostdc1", "Kcnj13","Calml4",
#Vascular endothelial cells", "venous
"Cldn5", "Ly6c1", "Flt1", "Ly6a", "Slco1a4", "Itm2a", "Cxcl12", "Bsg", "Pltp", "Ptprb",
#Vascular smooth muscle cells", "arterial
"Acta2", "Crip1", "Tagln", "Tpm2", "Myl9", "Myh11", "Igfbp7", "Myl6", "Vim", "Mylk",
#Ependymal cells
"Rarres2", "Dbi", "Tmem212", "Ccdc153", "Plac9b", "Fam183b", "Mia", "1700094D03Rik","Plac9a", "Dynlrb2",
#Red Blood Cells / Other
"Hba-a1", "Hbb-bs", "Hbb-bt", "Hba-a2", "Alas2", "Bpgm", "Isg20", "Ube2l6", "Tent5c","Rec114",
#Pericytes and Vascular leptomeningeal
"Ptgds", "Vtn", "Apod", "Mgp", "Dcn", "Ifitm1", "Igf2", "Col1a2", "Rgs5", "Nupr1",
"Cemip","Bicc1","Flrt2", "Lama2", # Perivascular_Fibroblast
#Neuronal intermediate progenitor cells
"Hist1h2ap", "Hmgb2", "Sox11", "Ccnd2", "Hmgn2", "Tubb2b", "Igfbpl1", "Sox4", "Dlx6os1","Ppp1r14b",
#Perivascular macrophages
"Lyz2", "Cd74", "Pf4", "Mrc1", "Selenop", "H2-Eb1", "C1qa", "H2-Ab1", "Fcer1g", "Stab1"
) %>% 
  unique()


scCustomize::Clustered_DotPlot(seurat_object = seurat, 
                               features = markers, cluster_feature = F, x_lab_rotate = 90,
                               group.by = "Level_4",  flip = T
                               ) 

```



```{r}
DotPlot(seurat, 
        features = markers, 
        group.by = "Level_1",
        cluster.idents = F 
        ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

DotPlot(seurat, 
        features = markers, 
        group.by = "Level_2",
        cluster.idents = F 
        ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


DotPlot(seurat, 
        features = markers, 
        group.by = "Level_3",
        cluster.idents = F 
        ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


DotPlot(seurat, 
        features = markers, 
        group.by = "Level_4",
        cluster.idents = F 
        ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```



[Chang, C., Zuo, H., & Li, Y. (2023). Recent advances in deciphering hippocampus complexity using single-cell transcriptomics. Neurobiology of Disease, 106062.](https://www.sciencedirect.com/science/article/pii/S0969996123000761)


```{r}
markers <- c(
  "Snap25", ## Synaptosomal Neuron
  "Grin2b", # Glutamate receptor Neuron
  "Neurod2", "Slc17a7", # Excitatory Neuron
  "Gad1", "Csf1r", # Inhibitory Neuron
  "Cx3cr1", "Aif-1", "Gfap", "Aqp4", "Aldh1l1",  # Microglia
  "Aldo", "Sox9", "Slc1a3", "Gjb6", "Mbp", "Plp1", "Mog", # Astrocyte
  "Mag", "Ermn", "Mobp", "Pdgfra", # Oligodendrocyte
  "Vcan", "Cspg4", "Olig2" # Oligodendrocyte precursor cell (OPC)
)

DotPlot(seurat, features = markers, cluster.idents = T, group.by = "Level_2") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

VlnPlot(seurat, features = markers, pt.size = 0,  stack = T)
dittoHeatmap(seurat, 
             markers[markers %in% rownames(seurat)],
             scale = "column", 
             cluster_rows = F,
             annot.by = c("Level_2", "condition"))

```


[Li Xuan, 2022](https://jneuroinflammation.biomedcentral.com/articles/10.1186/s12974-022-02447-5#Sec6)

```{r}
markers <- c(
"Cx3cr1", "Sall1", "P2ry12", "Fcrls", "Tmem119",
"Ccr2", "Fn1", "Cybb", "Pf4", "Ms4a7",
"Cxcr2", "S100a8", "Camp", "Ly6g",
"Flt3", "Cd209a", "Xcr1", "Ccr7", "Ccr9",
"Ncr1", "Klre1",
"Cd3e", "Lat", "Cd8b1", "Cd4",
"Cd19", "Ms4a1", "Cd79a",
"Gata3", "IL1rl1"
)

DotPlot(seurat, features = markers, cluster.idents = T, group.by = "Level_2") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

VlnPlot(seurat, features = markers, pt.size = 0,  stack = T)
dittoHeatmap(seurat, 
             markers[markers %in% rownames(seurat)],
             scale = "column", 
             cluster_rows = F,
             annot.by = c("RNA_snn_res.0.2", "condition"))

```


[Garza Raquel, 2023](https://www.cell.com/cell-reports/fulltext/S2211-1247(23)01407-9)

```{r}
markers <- c(
  "P2ry12", "Fyb1", "Itgax", " Ctss", " C1qb", "Cx3cr1",
  "Pmp2", "Col9a1", "Col6a1", "Vcan", "Pdgfra",
  "Plp1", "Mog", "Mbp",
  "Gfap", "Aqp4", "Gja1", "Slc1a3",
  "Bsg", "Flt1", "Lum", "Tpm1",
  "Dcx", "Rbfox1", "Grin1", "Hs3st2", "Gad1", "Gad2", "Calb2", "Cnr1",
  "Dcx", "Rbfox1", "Grin1", "Hs3st2"
) %>% 
  unique()

DotPlot(seurat, 
        features = markers, 
        group.by = "Level_2",
        cluster.idents = T 
        ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


VlnPlot(seurat, features = markers, pt.size = 0,  stack = T)
dittoHeatmap(seurat, 
             markers[markers %in% rownames(seurat)],
             scale = "column", 
             cluster_rows = F,
             annot.by = c("RNA_snn_res.0.2", "condition"))

```


Final plot for the paper

```{r}
markers <- c(
  "Grin2b", "Meg3", # excitatory neurons
  "Gad1", "Gad2", # inhibitory neurons
  "Mog", "Mbp", # oligodendrocytes
  "Gfap", "Slc1a2", # astrocytes
  "Cx3cr1" ,"C1qb", # microglia; 
  "Vcan", "Pdgfra", # oligodendrocyte precursor cells (OPC)
  "Sox11", "Ccnd2", # neuroblasts
  "Prlr", "Ttr", # choroid plexus epithelium
  "Tuba1a", "Ccdc153",  #  ependymal cells  
  "Flt1", "Pecam1", # endothelial
  "Flrt2", "Bicc1" # perivascular fibroblast 
) %>% 
  unique()
# paste0("italic('", gene_names, "')"), parse = T


scCustomize::Clustered_DotPlot(seurat_object = seurat, cluster_feature = F,
                               features = markers,
                               row_label_fontface = "italic",
                               x_lab_rotate = 90,
                               group.by = "Level_2", flip = F
                               ) 


scCustomize::Clustered_DotPlot(seurat_object = seurat, cluster_feature = F,
                               features = markers,  x_lab_rotate = 90,
                               row_label_fontface = "italic",
                               group.by = "Level_4", flip = F
                               ) 

```



